paper figures

### 2024-03-11 scRNAseq/paper_figures.Rmd
fixed IDs

- K456985_PBMC
- K783496_PBMC
- K693104_PBMC
- K955601_PBMC
- P1_1_PBMC
- P1_2_PBMC
- P3_1_PBMC
- P3_2_PBMC
- P8_1_PBMC
- P8_2_PBMC
- P22_1_PBMC
- P22_2_PBMC
- P23_1_PBMC
- P23_2_PBMC
- P25_1_PBMC
- P25_2_PBMC
- P27_1_PBMC
- P27_2_PBMC
- P28_1_PBMC
- P28_2_PBMC
- P22_P23_K783496_CD4
- P8_P3_K456985_CD4
- P27_P28_K693104_CD4
- P12_1_PBMC
- P12_2_PBMC
- P26_1_PBMC
- P26_2_PBMC
- K496398_PBMC
- P12_P26_K496398_CD4

@@

```{r get_metadata}
study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name))
assay <- read.csv(file.path(path,'sample_sheet','a_hegazy_IBD_totalseq_nucleotide_sequencing.txt'),
                 sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Library.Name=gsub('p','P',Library.Name))

@@

pbmc_meta <- do.call(rbind,
                     lapply(pbmc[grepl('PBMC',names(pbmc))], 
                            function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                             'orig.ident','sample',
                                                             'predicted.celltype.l1',
                                                             'predicted.celltype.l2',
                                                             'predicted.celltype.l3')) %>%
                              tibble::rownames_to_column('cell') %>%
                              dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

cols <- c('sample','orig.ident','predicted.celltype.l2')
all_frac <- do.call(rbind,lapply(pbmc, function(x) x[[cols]])) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map)) %>%
  dplyr::group_by(sample,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n)) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

pbmc_T_meta <- do.call(rbind,
                       lapply(pbmc.T,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
                                #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

```{r Fig_S4C,fig.width=6,fig.height=2}
cols <- c('sample','orig.ident','predicted.celltype.l2')
all_meta <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) x@meta.data[,cols])) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list()
for (ct in grep('CD[48]|Treg',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  tcrcomb <- lapply(tcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
  dfs[[ct]] <- clonalDiversity(tcrcomb[lapply(tcrcomb,function(x) dim(x)[1]) > 0],
                               cloneCall='gene',group='sample',exportTable=TRUE, skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint =ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','ITGB1','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

dfs <- list()
for (ct in c('Mmry CD4','Mmry CD8')) {
  bcs <- list(ITGA4=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4 > 0)],
              ITGB7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7 > 0)],
              Integrin_b7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='pos')],
              GPR15=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$GPR15 > 0)],
              CCR9=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$CCR9 > 0)],
              ITGAE=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGAE > 0)],
              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)],
              ITGB1=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1 > 0)],
              ITGA4neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4==0)],
              ITGB7neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7==0)],
              ITGB1neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1==0)],
              Integrin_b7_neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='neg')])
  for (nn in names(bcs)) {
    if (length(bcs[[nn]]) > 0) {
      tcrcomb <- lapply(tcr_combined,
                        function(x) x[(x$barcode %in% bcs[[nn]]) & !is.na(x$TCR1) & !is.na(x$TCR2),])
      dfs[[paste0(ct,'_',nn)]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
        dplyr::mutate(celltype=ct,
                      subset=nn)
    }
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

```{r Fig_S4D,fig.width=3,fig.height=2.5}
cols <- c('sample','orig.ident','predicted.celltype.l2')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=predicted.celltype.l2,
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

  dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  #dplyr::mutate(sample=ifelse(grepl('_CD4$',sample),sample,paste0(sample,'_PBMC'))) %>%
  dplyr::mutate(sample=paste0(sample,'_PBMC')) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=ifelse(Characteristics.Responder.=='yes','R','NR'),
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

```{r Fig_S4D_S5B,fig.width=4,fig.height=4}
dfs <- list()
for (ct in c('CD4 TCM','CD4 TEM',
             'CD8 TCM','CD8 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct) & (pbmc_md$Material=='PBMC')]
  if (length(bcs) > 0) {
    tcrcomb <- lapply(tcr_combined,
                      function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
    dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
      dplyr::mutate(celltype=gsub('CD[48] ','',ct),
                    Material=ifelse(grepl('CD4',ct),'CD4','CD8'))
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

pbmc_B_meta <- do.call(rbind,
                       lapply(pbmc.B,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
  #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                                   Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4')) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

```{r Fig_S5G,fig.width=4.5,fig.height=3}
dfs <- list()
for (ct in grep('B|Plasmablast',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  bcrcomb <- lapply(bcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !grepl('_NA|NA_',x$CTaa),])
  dfs[[ct]] <- clonalDiversity(bcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct,
                  Material='PBMC')
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::mutate(Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                                   Responder=Characteristics.Responder.,
                                   Diagnosis=Characteristics.Diagnosis.,
                                   Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                                   Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                                    ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl'))) %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

>>>>

- HC-1_PBMC
- HC-5_PBMC
- HC-2_PBMC
- HC-3_PBMC
- IBD-1_1_PBMC
- IBD-1_2_PBMC
- IBD-9_1_PBMC
- IBD-9_2_PBMC
- IBD-10_1_PBMC
- IBD-10_2_PBMC
- IBD-3_1_PBMC
- IBD-3_2_PBMC
- IBD-4_1_PBMC
- IBD-4_2_PBMC
- IBD-5_1_PBMC
- IBD-5_2_PBMC
- IBD-7_1_PBMC
- IBD-7_2_PBMC
- IBD-8_1_PBMC
- IBD-8_2_PBMC
- IBD-3_IBD-4_HC-5_CD4
- IBD-10_IBD-9_HC-1_CD4
- IBD-7_IBD-8_HC-2_CD4
- IBD-2_1_PBMC
- IBD-2_2_PBMC
- IBD-6_1_PBMC
- IBD-6_2_PBMC
- HC-4_PBMC
- IBD-2_IBD-6_HC-4_CD4

@@

```{r get_metadata}
# old IDs
study <- read.csv(file.path(path,'sample_sheet','s_hegazy_IBD.txt'),sep='\t') %>%
  dplyr::mutate(Sample.Name=gsub('p','P',Sample.Name),
                Patient.ID=gsub('[P]*([0-9]*)[\\/_][12]','P\\1',Characteristics.Patient.ID.),
                Responder=Characteristics.Responder.,
                Diagnosis=Characteristics.Diagnosis.,
                Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
                Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
                                 ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
                Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4'))

# new IDs
# study <- read.csv(file.path(path,'sample_sheet.txt'),sep='\t') %>%
#   dplyr::mutate(Patient.ID=Characteristics.Patient.ID.,
#                 Responder=Characteristics.Responder.,
#                 Diagnosis=Characteristics.Diagnosis.,
#                 Batch=gsub(' ','',gsub('_scRNAseq','',Parameter.Value.Method.)),
#                 Timepoint=ifelse(grepl('_1_',Sample.Name),'Vedo_0d',
#                                  ifelse(grepl('_2_',Sample.Name),'Vedo_6wk','ctrl')),
#                 Material=ifelse(Characteristics.Material.=='PBMC','PBMC','CD4'))

@@

pbmc_meta <- do.call(rbind,
                     lapply(pbmc[grepl('PBMC',names(pbmc))], 
                            function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                             'orig.ident','sample',
                                                             'predicted.celltype.l1',
                                                             'predicted.celltype.l2',
                                                             'predicted.celltype.l3')) %>%
                              tibble::rownames_to_column('cell') %>%
                              dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>%
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

cols <- c('sample','orig.ident','predicted.celltype.l2')
all_frac <- do.call(rbind,lapply(pbmc, function(x) x[[cols]])) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map)) %>%
  dplyr::group_by(sample,celltype) %>%
  dplyr::summarise(n=n()) %>%
  dplyr::mutate(frac=n/sum(n)) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

pbmc_T_meta <- do.call(rbind,
                       lapply(pbmc.T,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
                                #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

```{r Fig_S4C,fig.width=6,fig.height=2}
cols <- c('sample','orig.ident','predicted.celltype.l2')
all_meta <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) x@meta.data[,cols])) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

dfs <- list()
for (ct in grep('CD[48]|Treg',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  tcrcomb <- lapply(tcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
  dfs[[ct]] <- clonalDiversity(tcrcomb[lapply(tcrcomb,function(x) dim(x)[1]) > 0],
                               cloneCall='gene',group='sample',exportTable=TRUE, skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

cols <- c('sample','orig.ident','predicted.celltype.l2','ITGA4','ITGB7','ITGB1','Integrin-b7','GPR15','CCR9','ITGAE','MKI67')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map),
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

dfs <- list()
for (ct in c('Mmry CD4','Mmry CD8')) {
  bcs <- list(ITGA4=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4 > 0)],
              ITGB7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7 > 0)],
              Integrin_b7=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='pos')],
              GPR15=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$GPR15 > 0)],
              CCR9=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$CCR9 > 0)],
              ITGAE=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGAE > 0)],
              MKI67=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$MKI67 > 0)],
              ITGB1=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1 > 0)],
              ITGA4neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGA4==0)],
              ITGB7neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB7==0)],
              ITGB1neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$ITGB1==0)],
              Integrin_b7_neg=pbmc_md$cell[grepl(ct,pbmc_md$celltype) & (pbmc_md$Integrinb7=='neg')])
  for (nn in names(bcs)) {
    if (length(bcs[[nn]]) > 0) {
      tcrcomb <- lapply(tcr_combined,
                        function(x) x[(x$barcode %in% bcs[[nn]]) & !is.na(x$TCR1) & !is.na(x$TCR2),])
      dfs[[paste0(ct,'_',nn)]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
        dplyr::mutate(celltype=ct,
                      subset=nn)
    }
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

```{r Fig_S4D,fig.width=3,fig.height=2.5}
cols <- c('sample','orig.ident','predicted.celltype.l2')
pbmc_md <- do.call(rbind,lapply(pbmc[grepl('PBMC',names(pbmc))], function(x) FetchData(x,cols))) %>%
  tibble::rownames_to_column('cell') %>%
  dplyr::mutate(celltype=predicted.celltype.l2,
                cell=paste0(sample,'_',gsub('\\.','_',cell))) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name'))

@@

  dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct)
}
df <- do.call(rbind,dfs) %>%
  #dplyr::mutate(sample=ifelse(grepl('_CD4$',sample),sample,paste0(sample,'_PBMC'))) %>%
  dplyr::mutate(sample=paste0(sample,'_PBMC')) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

```{r Fig_S4D_S5B,fig.width=4,fig.height=4}
dfs <- list()
for (ct in c('CD4 TCM','CD4 TEM',
             'CD8 TCM','CD8 TEM')) {
  bcs <- pbmc_md$cell[(pbmc_md$celltype==ct) & (pbmc_md$Material=='PBMC')]
  if (length(bcs) > 0) {
    tcrcomb <- lapply(tcr_combined,
                      function(x) x[(x$barcode %in% bcs) & !is.na(x$TCR1) & !is.na(x$TCR2),])
    dfs[[ct]] <- clonalDiversity(tcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
      dplyr::mutate(celltype=gsub('CD[48] ','',ct),
                    Material=ifelse(grepl('CD4',ct),'CD4','CD8'))
  }
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)

@@

pbmc_B_meta <- do.call(rbind,
                       lapply(pbmc.B,#[grepl('PBMC',names(pbmc))], 
                              function(sobj) FetchData(sobj, c('refUMAP_1','refUMAP_2',
                                                               'orig.ident','sample',
                                                               'predicted.celltype.l2',
                                                               'CTaa',
                                                               'cloneType')) %>%
                                tibble::rownames_to_column('cell'))) %>%
  #dplyr::mutate(cell=paste0(orig.ident,'_',cell)))) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Material,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::mutate(celltype=plyr::revalue(predicted.celltype.l2,ct_map))

@@

```{r Fig_S5G,fig.width=4.5,fig.height=3}
dfs <- list()
for (ct in grep('B|Plasmablast',unique(all_meta$celltype),value=TRUE)) {
  bcs <- all_meta$cell[all_meta$celltype==ct]
  bcrcomb <- lapply(bcr_combined,
                    function(x) x[(x$barcode %in% bcs) & !grepl('_NA|NA_',x$CTaa),])
  dfs[[ct]] <- clonalDiversity(bcrcomb,cloneCall='gene',group='sample',exportTable=TRUE,skip.boots=TRUE) %>%
    dplyr::mutate(celltype=ct,
                  Material='PBMC')
}
df <- do.call(rbind,dfs) %>%
  dplyr::left_join(study %>% 
                     dplyr::select(Patient.ID,Responder,Diagnosis,Sample.Name,Timepoint,Batch),
                   by=c('sample'='Sample.Name')) %>%
  dplyr::arrange(Patient.ID)
